```{r, include = FALSE}
source(here::here("setup.R"))
for (file in list.files(here::here("R"), full.names = TRUE)) source(file)
```

```{r}
# List the files
list.files(data_dir)

# show the content of the labels file
# file.show(file.path(data_dir, "activity_labels.txt"))
cat(read_lines(file.path(data_dir, "activity_labels.txt")), sep = "\n")
```

The data files themselves are stored in the 'RawData' folder. This folder contains a folder 'Train' and a folder 'Test'. Let's look at the list of filenames in the 'Train' folder:

```{r}
# Train and Test folder are located in the RawData folder
list.files(file.path(data_dir, "RawData"))

# The Train folder contains 85 files:
length(list.files(train_dir))

# A sample of the file names in the Train folder; notice the filename pattern
set.seed(1)
sample(list.files(train_dir), 6)
```

# 1. Reading the data

* The format of signal files is `<sensor>_exp<XX>_user<YY>.txt`, where `<sensor>` is either `acc` (accelerometer) or `gyro` (gyroscope), `<XX>` is experiment ID, and `<YY>` is user ID.
* Each signal file contains three columns, one for each sensor channel, corresponding to the *X*, *Y*, and *Z* directions. The file consists of a sequence of samples.
* Instead of labeling every time point, the `labels_train.txt` file lists time intervals: each interval has a `start`, `end`, and activity code `act_code` (numbered 1 through 12).
* These intervals correspond to segments of the signals in `acc_expXX_userYY.txt` and `gyro_expXX_userYY.txt`.
* The names of activity codes are stored in the file `activity_labels.txt`.

**Goal:** To analyze the data, we need to merge the signals with their activity code and labels.
In practice, this means reading the signals (each row = one time point) and assigning the correct label to each time point.

## 1.1 Import Activity labels

```{r}
# Import the activity labels file
act_labels <- read.table(
  file.path(data_dir, "activity_labels.txt"),
  header = FALSE, col.names = c("code", "activity")
)

# Import the train labels file that contains the activity intervals for each trial and user
labels0 <- read_delim(file.path(train_dir, "labels_train.txt"),
  delim = " ",
  col_names = c("trial", "userid", "act_code", "start", "end"), col_types = "iiiii"
)

# Join the activity labels to the labels data frame
labels <- labels0 |> left_join(act_labels, by = c(act_code = "code"))

# Expand the labels data frame to have one row per sample
sample_labels <- labels |>
  mutate(segment = row_number()) |> # identify each separate activity interval
  reframe(
    sampleid = seq.int(start, end), # expand start:end into all sample IDs
    .by = c(trial, userid, act_code, activity, segment)
  )

# Check the result
cat("The sample_labels data frame has", nrow(sample_labels), "rows and", ncol(sample_labels), "columns.\n")
head(sample_labels)
tail(sample_labels)
```

## 1.2 Import the signals

To begin, we'll focus on a single file: the accelerometer signals for participant `user01` in trial `exp01`.

```{r}
username <- 1
expname  <- 1
sensor   <- "acc"
file_path_user01 <- get_file_path(train_dir, sensor, expname, username)
acc_signal_user01 <- load_signal(file_path_user01)
```

## 1.3 Merging Signals and Labels

We create a tibble that contains the `userid`, the `trial` number, and a `sampleid` for each row in `acc_signal_user01`.
Then we bind in the three signal columns (`X1`, `X2`, `X3`) and finally (left) join this table with `sample_labels`
so that every sample has the correct activity label.

```{r}
# Create a data frame with the signal and the corresponding activity labels
user_df <- get_signal_df(
  exp_id = expname,
  user_id = username,
  signal_data = acc_signal_user01,
  sample_labels = sample_labels
)

# Check the result (first few rows are not interesting; the following are)
cat("The user_df data frame has", nrow(user_df), "rows and", ncol(user_df), "columns.\n")
print(user_df[1227:1239, ])
```

This table is exactly what we need: each sample in the three signal channels is now labeled with its corresponding activity.

To check if the labeling is correct, we can visualize the result. The plot below shows the first signal channel (`X1`), with different colors indicating different activities:

```{r}
options(repr.plot.width = 15) # widen the plot for readability

# Plot the first signal channel, colored by activity
user_df %>%
  filter(!is.na(activity)) %>% # keep only labeled samples
  ggplot(aes(x = sampleid, y = X1, colour = factor(activity), group = segment)) +
  geom_line()
```

```{r, include = FALSE}
saveRDS(sample_labels, file.path(output_dir, "sample_labels.rds"))
saveRDS(user_df, file.path(output_dir, "user_df.rds"))
```
