```{r, include = FALSE}
source(here::here("setup.R"))
for (file in list.files(here::here("R"), full.names = TRUE)) source(file)
sample_labels <- readRDS(file.path(output_dir, "sample_labels.rds"))
user_df <- readRDS(file.path(output_dir, "user_df.rds"))
```

# 2. Feature Extraction

We continue working on signals from a single file, so we can develop iteratively without exhausting RAM.

We design our workflow so that we process one file at a time:
we extract features from it, store the smaller feature summary, and then move on to the next file.

## 2.1 Time domain features


### Looking at histograms

Histograms give a detailed picture of the distribution of the signal values.
For instance, have a look at the histograms of the signal samples per activity:

```{r}
user_df |>
  ggplot(aes(X1)) +
  geom_histogram(bins = 40, fill = 1, alpha = 0.5) +
  geom_histogram(aes(X2), bins = 40, fill = 2, alpha = 0.5) +
  geom_histogram(aes(X3), bins = 40, fill = 4, alpha = 0.5) +
  facet_wrap(~activity, scales = "free_y")
```

**The histograms for different activities look quite distinct.** This means that statistical summaries
(mean, standard deviation, skewness, etc.) could help us tell the activities apart.

* the center (mean),
* the spread (standard deviation, interquartile range, min-max range),
* the shape (skewness, kurtosis),
* ...and so on.

### Epochs


We also need an activity label for each epoch. Most epochs will have 128 samples with one and the same activity label.
However, because the partitioning of the signals into epochs of 128 samples is done without regard for the activity label,
some epochs will have more than one activity label, for example for a change in activity during that epoch.
We will deal with that by labeling such epochs with the label that is associated with the largest number of samples of that epoch.

Additionally, we will compute a confidence score for the activity label of each epoch,
which is the proportion of samples in that epoch that have the most common activity label.

### Writing your own feature functions


```{r}
# Compute features per epoch:
usertimedom <- get_time_domain_features(user_df)
head(usertimedom)

# Plot the confidence levels for each activity label
usertimedom |>
  filter(aggr_activity != "-") |>
  ggplot(aes(x = aggr_activity, y = activity_confidence)) +
  geom_boxplot() +
  facet_wrap(~aggr_activity, scales = "free_x") +
  theme(axis.text.x = element_blank()) +
  labs(x = "Activity label", y = "Confidence of activity label")
```

This is a tidy data frame that can be used in the classification algorithms discussed in chapters 3 and 4 of ISLR.

***Note**: You may not be familiar with the `AR1.1`, `AR1.2` and `AR12.1` computed above.
These are called _autocorrelations_ (at lags 1 and 2) and _lagged cross-correlations_.
They are discussed in [Feature Extraction from Signals](https://paper.dropbox.com/doc/Feature-extraction-from-Signals--A62tmtXDMS34X292NP0fKphQAQ-qCp5uvj47gmyuw5nmB8lL).
They are interesting because slowly changing signals have high autocorellations, while fast changing signals have low or even negative autocorrelations.*

## 2.2 Frequency domain features

In the previous section we extracted features from the **time domain** by looking at statistical summaries of the raw signals in fixed-length epochs.

We can also describe a signal in the **frequency domain**. Instead of asking "what values does the signal take?", we ask: *"how much of the signal is made up of slow changes vs fast changes?"*

A spectrum shows this: it tells us how much energy the signal has at different frequencies. You can think of it like analyzing a piece of music: the bass notes are low-frequency components, the melody is mid-frequency, and the hi-hats are high-frequency.

The steps are very similar to the time domain case:

1. **Split the signal into epochs** (same as before: 128 samples).
2. **Compute the spectrum for each epoch.**
3. **Summarise each spectrum** with a small number of frequency-domain features.

```{r}
# Step 1 + 2: compute spectra for each epoch
user_spect_df <- user_df |>
  mutate(epoch = sampleid %/% 128) |>
  reframe(
    {
      sp <- spectrum(cbind(X1, X2, X3), span = 15, plot = FALSE)
      tibble(
        freq = sp$freq, # cycles per sample
        spec1 = sp$spec[, 1],
        spec2 = sp$spec[, 2],
        spec3 = sp$spec[, 3]
      )
    },
    .by = epoch
  )

user_spect_df |> head()
```

Now we have a tidy data frame with the spectrum of each signal in each epoch.

Next, we summarise these spectra into features. The example spectral feature function `mean_frequency()` (defined in `R/`) computes the mean frequency from `freq` and `spec`. You can define your own.

```{r}
# Step 3: compute frequency-domain features per epoch
userfreqdom <- user_spect_df %>%
  group_by(epoch) %>%
  summarise(
    # Dominant frequency of signal 1
    domfreq1 = freq[which.max(spec1)],

    # Mean frequency of signal 2
    meanfreq2 = mean_frequency(freq, spec2),

    # Example of other possible features
    # (you can define your own functions like before)
    # spectral_entropy = ...
    # bandpower_low   = ...
    # bandpower_high  = ...
  )

head(userfreqdom)
```

### Recap

- In the time domain we used statistics like mean, sd, skewness per epoch.
- In the frequency domain we use statistics like dominant frequency, mean frequency, and energy in certain frequency bands.

The key idea is the same: **Each epoch becomes one row in a tidy data frame, with features that summarise the signal in that epoch.**

# 3. Putting it all together

We've just processed only one file. We need to process all of them and combine the resulting data frames in one big data frame. The function `extractTimeDomainFeatures()` (defined in `R/`) does this for the accelerometer data files (the ones starting with `acc`). You will need to do the same for the gyroscope data (filenames starting with `gyro`).

To obtain a list of all the files, use `dir()`:

```{r}
list.files(train_dir, pattern = "^acc")
```

This is not the most efficient way to organize a function (it combines several code chunks with very different functionality in one function; different types of functionality better go in separate functions), but it suffices for our purposes.

Note that the function is merely a copy and paste of the code developed above (with minor modifications). To see how it does as promissed:

```{r}
filename <- file.path(train_dir, "acc_exp01_user01.txt")
df <- extractTimeDomainFeatures(filename, sample_labels)
print(df)
```

It looks like our function works as expected (check this!).

Now, you could make a `for` loop to iterate through all the filenames listed by the `list.files()` command above, however, here's a slightly less cumbersome trick: Using `map_dfr()` that is part of the `tidyverse`. We demonstrate this with only 5 files; **you will have to remove that restriction**.

```{r}
# demonstrate this for only the first 5 files
filenames <- dir(train_dir, "^acc", full.names = TRUE)[1:10]

# map_dfr runs `extractTimeDomainFeatures` on all elements in
# filenames and binds results row wise
myData <- map_dfr(filenames, extractTimeDomainFeatures, sample_labels)

# Check the result (are there any NAs or unexpected values?)
summary(myData)
```

We really should check for NA's

Now the data are turned into a tidy format with features that can be used in any of the algorithms.

```{r, include = FALSE}
saveRDS(myData, file.path(output_dir, "myData.rds"))
```
