```{r, include = FALSE}
source(here::here("setup.R"))
for (file in list.files(here::here("R"), full.names = TRUE)) source(file)
user_df <- readRDS(file.path(output_dir, "user_df.rds"))
```

# Frequency domain features

## Spectrograms

As we looked at histograms of the signal values, we can also look at the spectragrams of the signals.
The spectrum is a representation of the signal in the frequency domain, which can reveal periodic patterns that are not obvious in the time domain.

```{r}
gaussian_sigma <- 3

# Plot the spectra for each activity on the log scale
user_df |>
  filter(!is.na(activity)) |>
  group_by(activity) |>
  reframe(
    spec1 = list(compute_spectrum(X1, gaussian_sigma = gaussian_sigma)),
    spec2 = list(compute_spectrum(X2, gaussian_sigma = gaussian_sigma)),
    spec3 = list(compute_spectrum(X3, gaussian_sigma = gaussian_sigma))
  ) |>
  ungroup() |>
  pivot_longer(cols = starts_with("spec"), names_to = "signal", values_to = "spectrum") |>
  unnest(cols = c(spectrum)) |>
  ggplot(aes(x = freq, y = spec, colour = signal)) +
  geom_line() +
  scale_y_log10() +
  facet_wrap(~activity) +
  labs(x = "Frequency (Hz)", y = "Spectral density (log scale)", colour = "Signal channel")
```

## Converting signals to the frequency domain

Describe a signal in the **frequency domain**.

```{r}
# Compute spectra for each epoch
user_spect_df <- convert_signal_to_spectrum_df(user_df)
user_spect_df |> head()
```

Now we have a tidy data frame with the spectrum of each signal in each epoch.

## Feature extraction

```{r}
# Compute frequency-domain features per epoch
userfreqdom <- get_frequency_domain_features(user_spect_df)
head(userfreqdom)
```

```{r, include = FALSE}
saveRDS(userfreqdom, file.path(output_dir, "userfreqdom.rds"))
```
